---
title: "Assignment 6 - Split-plot randomization"
author: "Ruben Vinueza"  
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: false
    theme: cerulean
---

# Introduction  
This assignment will ask you to conduct a split-plot RCBD randomization. 

For that, you will first need to watch (and code along with) the following split-plot randomization video:

- split-plot randomization: https://youtu.be/rXTp3Vw_770?si=MajBK1aZ221orfpa&t=3401 (starts at 00:56:41, watch till end, about 50-min total)

- The partial script for the video above is in our GitHub repository, 04-classcode, **02-13_spp_randomization_partial.qmd**.


**AFTER watching the video and developing the code in class**, then start working on this HW assignment.  

The different sections representing what we did in class were left in the script to provide you a scaffold to start from.  

# Instructions  
  - Move this script into the `code` subfolder of your `05_spp` project folder  
  - On this script, fill your name under the `title` section of the YAML  

# a) Setup  
Here is where we load the packages we will use.  
Make sure to load each package in the chunk below.  
```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)

```

# b) Treatment design  
# Split-plot in RCBD  
## WP randomization  
```{r wp_rcbd_rand}
#Setting seed to get some randomization result every time 
set.seed(930)

wp_rcbd_rand <- crossing(krate_kgha = c(0,30,60),
                         block = 1:4) %>%
  
arrange(block) %>%
# Randomizing Whole Plot treatment factor to blocks (restricted)
group_by(block) %>%
mutate(krate_kgha = sample (krate_kgha)) %>%
#Creating WP plot numbers
mutate(wplot = (block*100)+ 1:3) %>% 
  ungroup()

wp_rcbd_rand
```
How many rows do we have in the object above? Why?

## SP randomization  
Randomization of the split-plot treatment factor (N rate) to the WP treatment plots.
```{r sp_rcbd_rand}
sp_rcbd_rand <- crossing(nrate_kgha = c(0,100,200),
                         wplot = wp_rcbd_rand$wplot
                         ) %>%
  arrange(wplot) %>%
 # Randomizing SP treatment factor to WP plots (restricted)
 group_by(wplot) %>%
 mutate(nrate_kgha = sample(nrate_kgha)) %>%
 # Creating SP plot numbers
 mutate(splot = paste0(wplot,"-",1:3)) %>%
 ungroup()

sp_rcbd_rand
```

## Joining WP and SP into one object
```{r spp_rcbd_rand}
spp_rcbd_rand <- wp_rcbd_rand %>%
  right_join(sp_rcbd_rand) %>%
  mutate(trtname = paste0(krate_kgha, "+", nrate_kgha))

spp_rcbd_rand
```
How many rows in the object above? Why?

Inspecting block 1. 
```{r}
spp_rcbd_rand %>%
  filter(block ==1)
```
## Plot layout  
```{r spp_rcbd_layout}
spp_rcbd_layout <- spp_rcbd_rand %>%
  # Defining number of rows and columns in the field for plotting purposes
  mutate(row = rep(1:4, each = 9)) %>% 
  mutate(col = rep(seq(from = 1,
                       to = 9, #total number of treatments,
                       by = 1),
                   times = 4 #total number of blocks
                       )) %>%
  # Creating plot and trtname labels for plotting 
  mutate(lab = paste0(splot, "\n", trtname))

spp_rcbd_layout
```

```{r spp rcbd map}
ggplot(data = spp_rcbd_layout, aes(x = row,
                                   y = -col)) +
   geom_tile(aes(fill = factor(nrate_kgha)),
            color = "black",
            show.legend = F
            ) +
  geom_tile(aes(color = krate_kgha),
            show.legend = F,
            size = 2,
            linetype = 2,
            fill = NA
            ) +
  scale_fill_manual(values = c("#e5f0db",
                               "#b1d095",
                               "#5e813f")) +
  geom_text(aes(label = splot), fontface = "bold", vjust = -.8) +
  geom_text(aes(label = trtname), vjust = 1) +
  labs(title = "Split-plot RCBD randomization of a/n3 K (whole-plote) x 3 N (split-plot) treatment design") +
  theme_void()
 
```

# Exporting to file  

```{r}
spp_rcbd_rand %>%
  # Rearranging columns 
  dplyr::select(block, wplot, krate_kgha,
                splot, nrate_kgha,
                trtname) %>%
  write_csv("../output/spp_rcbd_randomization_table.csv")
```

```{r}
ggsave("../output/spp_rcbd_layoutmap.png",
       width = 6,
       height = 5,
       bg = "white"
       )
```
# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-06-spp-LASTNAME.html`.    

For ex., mine would be `Assignment-06-spp-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assigment #6 - Split-plot randomization** by Feb 20 11:59 pm.  
  





